Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-02-17 draft.tex
Updates from Overleaf

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_c.png}
    \caption{As per Fig. \ref{fig:rate_fitted} but with the data split by their colour parameter $c$.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_fitted_split_c}}
\end{figure}

@@

\subsection{Splitting by SN stretch \label{subsec:split_x1}}
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_x1_fitjoint.png}
    \caption{As per Fig. \ref{fig:rate_fitted_split_x1} but with the data fit by a two-component DTD.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_fitted_split_x1_2c_DTD}}
\end{figure}
The SN Ia rate as a function of stellar mass SNe split by stretch is shown in Fig. \ref{fig:rate_fitted_split_x1}. The evolution of the rate with stellar mass is significantly different for the two sub-samples: high-stretch SNe dominate the rate at low stellar mass but tail off in the higher mass galaxies, while low-stretch SNe are subdominant in low mass galaxies but display a much steeper dependence on stellar mass and make up the vast majority of SNe in high-mass hosts. These measurements reflect the observed correlations of $x_1$ with stellar mass and sSFR. The corresponding DTD fit results in a steep power-law $\beta = -1.68 \pm 0.15$ for high-stretch SNe indicative of a population of predominantly prompt SNe, and a much shallower decay slope ($\beta = -0.79 \pm 0.08$) for low-stretch SNe representing a much more delayed population. Such a distinction between SN stretch and progenitor age has been seen by \citet{Rigault2013,Childress2014,Rigault2018,Nicolas2020}; here we have shown it purely by measuring the rate against stellar mass.

@@

>>>>

@@

\subsection{Splitting by SN stretch \label{subsec:split_x1}}

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_x1_fitjoint.png}
    \caption{As per Fig. \ref{fig:rate_fitted_split_x1} but with the data fit by a two-component DTD.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_fitted_split_x1_2c_DTD}}
\end{figure}
The SN Ia rate as a function of stellar mass SNe split by stretch is shown in Fig. \ref{fig:rate_fitted_split_x1}. The evolution of the rate with stellar mass is significantly different for the two sub-samples: high-stretch SNe dominate the rate at low stellar mass but tail off in the higher mass galaxies, while low-stretch SNe are subdominant in low mass galaxies but display a much steeper dependence on stellar mass and make up the vast majority of SNe in high-mass hosts. These measurements reflect the observed correlations of $x_1$ with stellar mass and sSFR. The corresponding DTD fit results in a steep power-law $\beta = -1.68 \pm 0.15$ for high-stretch SNe indicative of a population of predominantly prompt SNe, and a much shallower decay slope ($\beta = -0.79 \pm 0.08$) for low-stretch SNe representing a much more delayed population. Such a distinction between SN stretch and progenitor age has been seen by \citet{Rigault2013,Childress2014,Rigault2018,Nicolas2020}; here we have shown it purely by measuring the rate against stellar mass.

@@

\subsection{Splitting by SN colour \label{subsec:split_c}}
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_c.png}
    \caption{As per Fig. \ref{fig:rate_fitted_split_x1} but with the data split by their colour parameter $c$.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_fitted_split_c}}
\end{figure}

In the previous section, we have showed that the slope of the SN Ia rate vs stellar mass relation is different for slow and fast declining supernovae. The same is for the most part not true for SNe with different colours (Fig. \ref{fig:rate_fitted_split_c}), with best fitting values of $\beta$ of $-1.19\pm0.09$ $-1.09\pm0.08$ for red ($c > 0$) and blue ($c<0$) SNe respectively. This lack of difference is in agreement with the findings of \citet{Sullivan2010}. Red SNe appear to have a higher normalisation than blue SNe, with roughly 30\% more red than blue SNe. This dominance can be explained by a single population of SNe with an intrinsic colour distribution that is Gaussian and centred on $c=0$ combined with an external dust distribution with an exponential form, systematically shifting the SN colour distribution to the red and causing the higher observed rate of those objects. Such a scenario has been proposed by \citet{Brout2020}.

The predictions from the DTD model diverge from the data at low stellar mass ($M_* <10^{9.25}$ M$_{\odot}$), where the rate becomes dominated by blue SNe. Although this is where the data is noisiest, it is worth investigating the possible causes of this divergence. A slight preference for SN to be bluer in lower-mass galaxies has been observed \citep{Scolnic2018,Brout2019,Smith2020,Kelsey2020} and could be the same effect seen at low stellar masses here. Physically, the dominance of blue SNe at low stellar masses could either be an intrinsic property of the SNe such as a hotter photosphere in explosions of younger, more massive progenitors, or an extrinsic effect such as a lower average dust column density in low-mass host galaxies. An intermediate explanation is that the SNe in low-mass hosts have a smaller dust column density in the immediate vicinity of the progenitor system, either due to the shorter evolutionary timescale or the higher mass and thus brighter system being less conducive to dust production. A detailed exploration of the effects of environment on SN colour will be presented in Kelsey et al. \textit{in prep}.
